---
title: "Untitled"
author: "Haley Oetterer"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
#in this there is asprin use vs cacner diagnosis with age and sex not confounded for. 
#Aspirin vs cancer not using sex or age at all. 
#Aspirin vs Cancer with both as binary variables. 
#Aspirin vs Cancer with age and sex as confounders 
#Aspriin vs cancer with only age as confouder excluding sex
```

```



```{r}
# Load libraries
library(dplyr)
library(ggplot2)


# Define the maximum number of rows based on your data
n <- 9254  # Total rows from Age count

# Create the dataset
data <- data.frame(
  Age = sample(c("<20", "20-39", "40-59", "60+"), size = n, replace = TRUE, prob = c(121, 3411, 3364, 2358) / 9254),
  Sex = sample(c("male", "female"), size = n, replace = TRUE, prob = c(4629, 4625) / 9251),
  Cancer = sample(c(1, 0), size = n, replace = TRUE, prob = c(46, 8851) / 8897),
  Aspirin_Use = sample(c("daily", "every_other_day", "another_schedule", "none"), size = n, replace = TRUE, prob = c(1034, 69, 122, 2657) / 3882)
)

```



```{r}
#table1 
install.packages("tableone")
library(tableone)

# Define categorical variables
categorical_vars <- c("Age", "Sex", "Cancer", "Aspirin_Use")

# Create the Table 1
table1 <- CreateTableOne(vars = categorical_vars, data = data, factorVars = categorical_vars)

# Print Table 1
print(table1)


```




```{r}
# Fit logistic regression
model <- glm(Cancer ~ Aspirin_Use + Age + Sex, family = binomial(link = "logit"), data = data)

# Summary of the model
summary(model)

# Odds Ratios and Confidence Intervals
exp(cbind(OddsRatio = coef(model), confint(model)))

```


```{r}
library(ggplot2)

# Predict probabilities
data$predicted <- predict(model, type = "response")

# Plot
ggplot(data, aes(x = Aspirin_Use, y = predicted, color = Age)) +
  geom_point() +
  facet_wrap(~Sex) +
  labs(title = "Predicted Cancer Risk by Aspirin Use",
       x = "Aspirin Use",
       y = "Predicted Probability of Cancer") +
  theme_minimal()
```


```{r}
#only aspirin use vs cancer resgression line graph steps 
n <- 9254

# dataset
data <- data.frame(
  Aspirin_Use = sample(c("daily", "every_other_day", "another_schedule", "none"), size = n, replace = TRUE, 
                       prob = c(1034, 69, 122, 2657) / 3882),
  Cancer = sample(c(1, 0), size = n, replace = TRUE, prob = c(46, 8851) / 8897)
)

# Fit logistic regression model
model <- glm(Cancer ~ Aspirin_Use, family = binomial(link = "logit"), data = data)

# View model summary
summary(model)

# Odds Ratios and Confidence Intervals
exp(cbind(OddsRatio = coef(model), confint(model)))

```


```{r}
#line graph aspirinuse vs cancer
# prediction
prediction_data <- data.frame(Aspirin_Use = unique(data$Aspirin_Use))
 library(ggplot2)

# Plot the logistic regression line with adjusted axis limits
ggplot(data, aes(x = Aspirin_Use, y = Cancer)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), color = "blue", se = FALSE) +
  geom_point(aes(y = as.numeric(Cancer)), position = position_jitter(width = 0.2, height = 0.02), alpha = 0.4) +
  scale_y_continuous(limits = c(0, 0.1)) +  # Adjust the y-axis to zoom in
  labs(title = "Logistic Regression: Aspirin Use and Cancer (Zoomed In)",
       x = "Aspirin Use",
       y = "Probability of Cancer") +
  theme_minimal()

```


```{r}
#cancer vs aspirin use, both binary vairables
# Create a binary variable for aspirin use
data$Aspirin_Use_Binary <- ifelse(data$Aspirin_Use == "none", "No", "Yes")

# Fit logistic regression model
model_binary <- glm(Cancer ~ Aspirin_Use_Binary, family = binomial(link = "logit"), data = data)

# View model summary
summary(model_binary)

# Create dataset for prediction
prediction_data_binary <- data.frame(Aspirin_Use_Binary = c("No", "Yes"))

# Predict probabilities based on the model
prediction_data_binary$predicted_probability <- predict(model_binary, newdata = prediction_data_binary, type = "response")

# Odds Ratios and Confidence Intervals
exp(cbind(OddsRatio = coef(model), confint(model)))


```

```{r}
#graph of binary cancer vs aspirin use 

library(ggplot2)

# Plot the logistic regression
ggplot(prediction_data_binary, aes(x = Aspirin_Use_Binary, y = predicted_probability)) +
  geom_bar(stat = "identity", fill = "blue", width = 0.5) +
  labs(title = "Logistic Regression: Cancer Diagnosis vs Aspirin Use (Binary)",
       x = "Aspirin Use (Yes/No)",
       y = "Predicted Probability of Cancer") +
  theme_minimal()

```


```{r}
#sex & age as confounders ---------------------------

# Define the total number of rows based on the largest count (Age: 9254)
n <- 9254

# Simulate the dataset
data <- data.frame(
  Age = sample(c("<20", "20-39", "40-59", "60+"), size = n, replace = TRUE, prob = c(121, 3411, 3364, 2358) / 9254),
  Sex = sample(c("male", "female"), size = n, replace = TRUE, prob = c(4629, 4625) / 9251),
  Cancer = sample(c(1, 0), size = n, replace = TRUE, prob = c(46, 8851) / 8897),
  Aspirin_Use = sample(c("daily", "every_other_day", "another_schedule", "none"), size = n, replace = TRUE, prob = c(1034, 69, 122, 2657) / 3882)
)


```


```{r}
# Fit logistic regression model with Age and Sex as confounders
model <- glm(Cancer ~ Aspirin_Use + Age + Sex, family = binomial(link = "logit"), data = data)

# Summary of the model
summary(model)

# Compute Confidence Intervals
confint(model)  

# Odds Ratios and Confidence Intervals
exp(cbind(OddsRatio = coef(model), confint(model)))

```


```{r}
library(ggplot2)

# Predict probabilities using the fitted model
data$predicted <- predict(model, type = "response")

# Plot the predicted probabilities
ggplot(data, aes(x = Aspirin_Use, y = predicted, color = Age)) +
  geom_point(position = position_jitter(width = 0.2, height = 0.01)) +
  facet_wrap(~Sex) +
  labs(title = "Predicted Cancer Risk by Aspirin Use (Adjusted for Age and Sex)",
       x = "Aspirin Use",
       y = "Predicted Probability of Cancer") +
  theme_minimal()

```

```{r}
#just age as confouder excluding sex
# Define the total number of rows based on the largest count (Age: 9254)
n <- 9254

# Simulate the dataset
data <- data.frame(
  Age = sample(c("<20", "20-39", "40-59", "60+"), size = n, replace = TRUE, prob = c(121, 3411, 3364, 2358) / 9254),
  Cancer = sample(c(1, 0), size = n, replace = TRUE, prob = c(46, 8851) / 8897),
  Aspirin_Use = sample(c("daily", "every_other_day", "another_schedule", "none"), size = n, replace = TRUE, prob = c(1034, 69, 122, 2657) / 3882)
)

```


```{r}
# Fit logistic regression model with Age as a confounder
model <- glm(Cancer ~ Aspirin_Use + Age, family = binomial(link = "logit"), data = data)

# Summary of the model
summary(model)

```

```{r}
library(ggplot2)

# Predict probabilities
data$predicted <- predict(model, type = "response")



# Plot the predicted probabilities
ggplot(data, aes(x = Aspirin_Use, y = predicted, color = Age)) +
  geom_point(position = position_jitter(width = 0.2, height = 0.01)) +
  labs(title = "Predicted Cancer Risk by Aspirin Use (Adjusted for Age)",
       x = "Aspirin Use",
       y = "Predicted Probability of Cancer") +
  theme_minimal()

```


```{r}



```









```{r}
